---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: true
---


## Loading and preprocessing the data

We download, extract and import the data.
```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",
              "repdata-data-activity.zip")
unzip("repdata-data-activity.zip")
activity <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?

The first assignment is divided in 2 parts. The first is
 
 > Make a histogram of the total number of steps taken each day

This is done with a `barplot`
```{r}
totalstepsperday <- tapply(activity$steps, activity$date, sum, na.rm = TRUE)
barplot(totalstepsperday, xlab = "Days", ylab = "Total steps taken")
```

The second part is 

 > Calculate and report the mean and median total number of steps taken
per day

This is done for the mean by
```{r}
tapply(activity$steps, activity$date, mean, na.rm = TRUE)
```
and for the median by
```{r}
tapply(activity$steps, activity$date, median, na.rm = TRUE)
```
The data look really bad. Not so many `NA`s, but a lot of `0`s.
This makes some computations useless, like the median that is always `0`.

## What is the average daily activity pattern?

Again, two main steps for this assignment

 > Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all days (y-axis)

First the plot, again with `base` plotting system

```{r}
meanstepsperinterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(unique(activity$interval), meanstepsperinterval, type = "l", xlab = "Interval number",
     ylab = "Average of steps")
```

 > Which 5-minute interval, on average across all the days in the dataset,
contains the maximum number of steps?

Second the most used, in average, time frame to walk

```{r}
names(meanstepsperinterval)[which(meanstepsperinterval %in% max(meanstepsperinterval))]
```

## Imputing missing values
This assignment has `4` steps

 > Calculate and report the total number of missing values in the dataset
 (i.e. the total number of rows with NAs)

There are many `NA`s; more than 10% of the dataset is filled with `NA`s.
```{r}
sum(is.na(activity$steps))
sum(is.na(activity$steps))/length(activity$steps)*100
```


 > Devise a strategy for filling in all of the missing values in the dataset.
 The strategy does not need to be sophisticated. 
 For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

I decide to fill the `NA`s with the mean of the steps per interval

 > Create a new dataset that is equal to the original dataset but with the missing data filled in.

The `newactivity` dataset is created here

```{r}
newactivity <- activity

for (i in 1:dim(newactivity)[1]) {
    if(is.na(newactivity[i,1])) {
        newactivity[i, 1] <- meanstepsperinterval[as.character(newactivity[i,3])]
    }
}
```

Last part of this assignment is

 > Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
 Do these values differ from the estimates from the first part of the assignment?
 What is the impact of imputing missing data on the estimates of the total
daily number of steps?

Again with `base` plotting, the histrogram is a `barplot`

```{r}
newtotalstepsperday <- tapply(newactivity$steps, newactivity$date, sum, na.rm = TRUE)
barplot(newtotalstepsperday, xlab = "Days", ylab = "Total steps taken - filled database")
```

And these are mean and median of this new dataset

```{r}
tapply(newactivity$steps, activity$date, mean, na.rm = TRUE)
tapply(newactivity$steps, activity$date, median, na.rm = TRUE)
```

We can see that by adding these missing datas we did not touch the datas that we 
had before already. In particular the median is still not significative, because
it is due to too many `0`s in the dataset.  
And for the mean, we just mutated the `0`s in a positive value.

## Are there differences in activity patterns between weekdays and weekends?
Last two assignments of the exercise

> Create a new factor variable in the dataset with two levels – “weekday”
and “weekend” indicating whether a given date is a weekday or weekend
day.

```{r}

weekends <- weekdays(as.Date(newactivity$date), abbreviate = T) == "Sun" |
    weekdays(as.Date(newactivity$date), abbreviate = T) == "Sat"
newactivity$weekdays <- factor(weekends, labels = c("weekday", "weekend"))
```


 > Make a panel plot containing a time series plot (i.e. type = "l") of the
5-minute interval (x-axis) and the average number of steps taken, averaged
across all weekday days or weekend days (y-axis). The plot should look
something like the following, which was creating using simulated data:

I do this last assignment with the `lattice` plotting system.
```{r}
library(lattice) 
pl <- xyplot(steps ~ interval | weekdays, type = "l", data = newactivity,
             layout=c(1,2), xlab = "Interval number", ylab = "Number of steps")
print(pl)
```